---
title: "kwpower R.package"
author: "Ching Yao Yang"
date: "25 April 2023"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# kwpower R.package

The **kwpower package** allows for the efficient comparison of mean ranks between three or more groups with continuous but not normally distributed data by performing the **Kruskal-Wallis Test** and **Multiple Pairwise-Comparison.** This eliminates the need to run these tests separately, saving time and effort. Additionally, the kwpower package produces graphics that visualize the data, making it easier to interpret the results.

The package's main function is called **kwp()**, and it is accompanied by two other functions, **ppt1()** and **ppt2()**. These functions generate plots that facilitate easy comparison. Users can input their dataset, and the package will automatically produce the plots.

For example, we would like to test whether the **average monthly temperatures in Tampa in 1900, 1990,** **and 2022** have been significantly affected by global warming, and visualize the data accordingly. To do this, we will use the dataset "1900_1990_2022_Tampa_monthly_temperature.csv".

One of the **key features** of the kwpower package is its ability to test **any type of CSV data** and generate plots with a structure that includes values in the **2nd column** and groups in the **3rd column**. To read a CSV file, users can choose the file interactively using a dialog box.

**"With just a few clicks, users can perform their analysis easily."**

### **Dataset**

I use a dataset called 1900 1990 2022 Tampa monthly temperature.csv in the data folder to test whether is any significant difference between groups which in here are 1900 1990 2022 and the values of groups are monthly temperature.

In this Markdown, we use the data frame for easy access to the data without having to read the file.

```{r}
my_data<- df <- data.frame(
  month = rep(c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"), 3),
  temperature = c(40.1, 46.3, 53.3, 58.9, 64.9, 71.3, 73.3, 72.4, 68.8, 62.4, 56, 56,
                  65.9, 69, 69.5, 72, 80.4, 82.6, 82.4, 83.7, 82.6, 77.5, 70.1, 66.8,
                  61.2, 66.9, 71.9, 75.6, 80.8, 83.9, 84.8, 84.8, 82.5, 76, 73.5, 65.5),
  year = c(rep(1900, 12), rep(1990, 12), rep(2022, 12))
)
my_data
```

### **1. kwp() function**

1.The function checks whether the **"dplyr" package** is installed. If it is not, the function installs it using the "install.packages" function. It then loads the package using the "library" function.

2.The function performs the Kruskal-Wallis test to compare the variable across the different groups. The resulting test statistic, degrees of freedom, and ***p*****-value** are saved in the kwp variable.

3.The function checks the p-value of the Kruskal-Wallis test and prints a message depending on whether the ***p-value*** is greater than or equal to 0.05. If the ***p-value*** is greater than or equal to 0.05, it prints **"No significance between groups, try with other dataset"** and stops the execution.

4.If the ***p-value*** is less than 0.05, the function prints **"Observe Significance between groups, continue pairwise_wilcox comparisons"**. It then adjusts the *p-values* using the Benjamini-Hochberg method and saves the results of the pairwise Wilcoxon tests in the pairwise_wilcox variable. The results are printed.

```{r}
kwp <- function(file_path = NULL, xvar = NULL, yvar = NULL) {
  
  # Check if dplyr package is installed, if not install it
  if (!require(dplyr)) {
    install.packages("dplyr")
    library(dplyr)
  }
  
  # Check if xvar and yvar are specified, if not, use the second and third columns
  if (is.null(xvar) & is.null(yvar)) {
    xvar <- names(my_data)[3]
    yvar <- names(my_data)[2]
  }
  
  # Check levels of xvar column
  my_data[[xvar]] <- factor(my_data[[xvar]])
  cat("Levels of", xvar, "variable: ", levels(my_data[[xvar]]), "\n")
  
  # Perform Kruskal-Wallis test
  test_result <- kruskal.test(reformulate(xvar, yvar), data = my_data)
  
  if (test_result$p.value >= 0.05) {
    message("\'~~No significance between groups, try with other dataset~~\'")
  } else {
    message("\'~~Observe Significance between groups, continue pairwise_wilcox comparisons~~\'")
    
    pairwise_wilcox <- my_data %>%
      select({{yvar}}, {{xvar}}) %>%
      {pairwise.wilcox.test(.[[1]], .[[2]], p.adjust.method = "BH", by = .[[2]])}
    
    pairwise_wilcox
  }
}
kwp(my_data)
```

### 2.ppt1() function

1.This function creates a **"special plot"** that displays the **data distribution in a line** for each group, making it easy to compare the data distributions across different groups.

2.The function checks for the **ggplot2 package** and installs it automatically if it is not already installed on the system.

3.When running the function, a file selection **dialog box** will appear. Users should select a CSV file where the values are located in the 2nd column and the groups are in the 3rd column. In this Markdown, we use the data frame for easy access to the data without having to read the file.

4.The function creates a scatter plot (using geom_point()) with loess smoothing (using geom_smooth()) for each group and its corresponding values, and saves the resulting plots in a variable called ppt1.

```{r}
ppt1 <- function(my_data) {
  # Check if ggplot2 package is installed, if not install it
  if (!require("ggplot2")) {
    install.packages("ggplot2")
    library("ggplot2")
  }
  
  # Create plot
  group_col <- names(my_data)[3]
  weight_col <- names(my_data)[2]
  my_data[[group_col]] <- as.factor(my_data[[group_col]])
  i <- ggplot(my_data, aes(x = !!sym(group_col), y = !!sym(weight_col), color = !!sym(group_col))) +
    geom_point() +
    geom_smooth(method = "loess") +
    scale_color_viridis_d() +
    theme(legend.position = c(1.5, .8))+
    ggtitle("Data by group")+
    theme(plot.title = element_text(size = 12, color = "red", hjust = 0.5))
  plot(i)
}
ppt1(my_data)
```

### 3.ppt2() function

1.This function creates a **"box plot"** that makes it easy to compare the data distribution across different groups. A box plot displays the **median**, **quartiles**, and **outliers** of each group's data.

2.The function checks if the **ggpubr package** is installed and installs it automatically if it is not already installed on the system. The ggpubr package provides easy-to-use functions for creating high-quality publication-ready graphs.

3.When running the function, a file selection **dialog box** will appear. Users should select a CSV file where the values are located in the 2nd column and the groups are in the 3rd column. In this Markdown, we use the data frame for easy access to the data without having to read the file.

4.The function creates a box plot using the data and saves the resulting plot in a variable called "boxplot". This plot can be further customized using ggplot2 functions.

```{r}
ppt2 <- function(my_data, group_col = NULL, weight_col = NULL) {
  # Check if ggpubr package is installed, if not install it
  if (!require("ggpubr")) {
    install.packages("ggpubr")
    library("ggpubr")
  }
 
  if (is.null(group_col)) {
    group_col <- names(my_data)[3]
  }
  if (is.null(weight_col)) {
    weight_col <- names(my_data)[2]
  }
  group_levels <- unique(my_data[[group_col]])
  palette <- viridisLite::viridis(length(group_levels))
  names(palette) <- group_levels
  j <- ggboxplot(my_data, x = group_col, y = weight_col,
                 color = group_col, palette = palette,
                 ylab = weight_col, xlab = group_col)
  plot(j)
}
ppt2(my_data)
```
